Notes on Perturbational Quantities

### 2008-07-30 docs/pert_notes/paper.tex
more details

We define $\rho'^n = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 

@@

We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves 
(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the 
superscript indicates that we have not called correct\_base yet (correct\_base takes 
$\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ as the output).  We compute $\rho$ on 
edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.  We 
use $\rho^{(2)}$ rather than $\rho^{n+1}$ to be consistent with the fact that we did not 
include an $\eta_\rho$ contribution to compute $\rho'$ at edges.  For the non-radial edges, 
we can directly add $\rho_0^{n+\half}$ to $\rho'$ since $\rho_0^{n+\half}$ is a cell-centered 
quantity.  For the radial edges, we interpolate to obtain $\rho_0^{n+\half}$ at radial edges 
before adding it to $\rho'$.
\subsection{Predicting $\rho$ (predict\_rho = T)}
\subsubsection{Predicting $\rho$ at edges}
We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term 
is set to zero, but then we set is\_conservative = T to use the conservative 
discretization in make\_edge\_scal.
\subsection{Predicting $\rho_0$ (predict\_rho = T or F)}
To advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be used in 
correct\_base, we need $\rho_0$ at edge-centered, time-centered locations.  However, we do 
not use the same edge-centered, time-centered $\rho_0^{n+\half}$ referred to in Section 
\ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed with a Godunov method in 
make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the term proportional to 
$\rho_0$ in (\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.

@@

@@

+ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"} \label{rhohprime equation}

@@

$(\rho h)'^{(1)} = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'^{(1)}$ to 

@@

(\ref{rhohprime equation}) as the 
forcing.

@@

\subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}
\subsection{Predicting $T$ into $(\rho h)'$ (enthalpy\_pred\_type = 3)}
\subsection{Predicting $T$ into $h$ (enthalpy\_pred\_type = 4)}

>>>>

We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 

@@

\subsubsection{Predicting $\rho_0$ at edges}
There are two ways to predict $\rho_0$ at edges.\\ \\
We can call make\_edge\_state\_1d using the term proportional to $\rho_0$ in 
(\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.

@@

This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be 
used in correct\_base.\\ \\
We can also define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We first 
compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves (\ref{rho0 equation}) 
through $\Delta t$ in time.  The $(2)$ in the superscript indicates that we have not called 
correct\_base yet (correct\_base takes $\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ 
as the output).  This term is used to construct $\rho$ at edges from $\rho'$ at edges, and 
this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  We use $\rho^{(2)}$ rather 
than $\rho^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be consistent with the fact 
that we do not include an $\eta_\rho$ contribution to compute $\rho'$ at edges.
For the non-radial edges, we can directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 
$\rho_0^{n+\half,{\rm avg}}$ is a cell-centered quantity.  For the radial edges, we 
interpolate to obtain $\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to $\rho'$.
\subsubsection{Advancing $\rho X_k$}
The evolution equation for $\rho X_k$, ignoring the reaction terms that were already 
accounted for in react\_state, and the associated discretization is
\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}}+\rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.

@@

+ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}

@@

$(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 

@@

(\ref{rhohprime equation}) as the forcing.\\ \\
{\bf What the code is currently doing:} Instead of defining 
$(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries through the complete 
evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ react\_base 
$\rightarrow (\rho h)_0^{(1)} \rightarrow$ advect\_base 
$\rightarrow (\rho h)_0^{(2)} \rightarrow$ react\_base $\rightarrow (\rho h)_0^{n+1}$.  
There is no correct\_base function for $(\rho h)_0$ written yet, nor is any adjustment made 
to sync up $(\rho h)_0$ after the thermal diffusion setp.
\subsubsection{Predicting $(\rho h)_0$ at edges}

@@

\subsubsection{Advancing $\rho h$}
We would like to update the non-forcing part of the enthalpy equation using
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = -\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}}+(\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\}.
{\bf What the code is currently doing:}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt + w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\right].
